Introduction to\\ Swedish Civil Contingencies Agency (MSB) methodological support for introducing \ac{isms}.

### 2019-04-01 msbframework/msbintro/msbintro.tex
Changed the order of governance documents and implementing an ISMS
Fixed some small spelling misstakes

\begin{frame}{Tieto incident}
  \begin{itemize}
    \item In 2011 Tieto got a hardware problem in one of their
      datacenters~\cite{Lindkvist2012tdf}. This created disruptions in many of
      the companies customers IT-system. For example: 
      \begin{itemize}
        \item All the pharmacies got problems handing out prescription medicin for a
          time.
        \item The swedish vehivle inspection agency could not report the results
          from the vehicle inspections they did during this time.
        \item A large logistical company could not handle its operation
        \item Social offices in Nacka and Sollentuna were unable to pay child
          support
      \end{itemize}
  \end{itemize}
\end{frame}
\begin{frame}{1177 incident}
  \begin{itemize}
    \item In 2019 a sub-contractor of a sub-contractor for a company that was
      handling the IT-system for three swedish regions health care guides (1177)
      call line, forgot to secure a backup server that stored the recordings
      from the phone calls~\cite{Ottsjo2019}.
    \item Resulted in around 2.7 million (around 170 000 hours) of recordings of
      confidential phone calls got accessible for the public.
    \item The recordings contains everything from medical symptoms to social
      security numbers.
  \end{itemize}
\end{frame}
\begin{frame}{The Swedish Transport Agency}
  \begin{itemize}
    \item In 2017 it was revealed that sensitive information from the Swedish
      Transport Agency database were made available for non-security screened
      employees abroad~\cite{SvTTransport2017}. 
    \item Information included:
      \begin{itemize}
        \item Driver licences (including picture)
        \item Company secrets
        \item Sensitive information about infrastructure (bridges, subways etc)
        \item Information that can be used to identify armoured cars.
      \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Swedish Public Employment Service}
  \begin{itemize}
    \item In 2019 it was found that the Swedish Publich Employment Services
      (arbetsförmedlingen) automatic system for reporting activity reports to
      the Swedish Social Insurance Agency (Försäkringskassan) sent incorrect
      reports~\cite{SvTPublicEmpl2019}.
    \item It was estimated that 10 to 15 percent of the decisions were incorrect.
    \item Resulted in that approximately 70 000 decisions could have been wrong.
  \end{itemize}
\end{frame}

@@

\begin{frame}{Dividing informational assets}{\insertsubsubsectionhead}
  \begin{description}
    \item[Primary] Refers to the main information, such as blue prints, logs or
      contracts.
    \item[Secondary] Refers to resources that are required to access the primary
      informational assets.
  \end{description}
  What will happen if you do not have your program that is required to read the
  closed proprietary format that the information is stored with?
\end{frame}
\begin{frame}{Critical Informational Assets}
  \centering
  \begin{block}{What is a critical informational asset?}
    There are alot of informational assets in an organisation. It is therefore
    necessary to establish what should be considered a critical asset.
  \end{block}
\end{frame}

@@

\subsubsection{Risk analysis}
\begin{frame}{Overview}{\insertsubsubsectionhead}
  \begin{itemize}
    \item Identifies the information security risks.
    \item Can be performed over generally over the organisation, or on a
      specific object.
    \item The identified risks are then assessed based on probability and
      consequence.
  \end{itemize}
  \begin{block}{What to acheive with a risk analysis}
    \begin{itemize}
      \item What can happen?
      \item How probable is it?
      \item What is the consequences?
    \end{itemize}
    Note that the result of the risk analysis shouldn't be seen as the absolute
    truth.
  \end{block}
\end{frame}

\begin{frame}{When should the Risk Analysis be performed?}{\insertsubsubsectionhead}
  \begin{itemize}
    \item Not necessary to perform the risk analysis after analysing the
      organisation or external and legal requirements.
    \item If it is performed after, then the results from these can be used in
      the risk analysis.
    \item Otherwise, it is possible to rely on the existing knowledge of the
      organisation.
  \end{itemize}
\end{frame}
\begin{frame}{When should the Risk Analysis be performed?}{\insertsubsubsectionhead}
  Examples of when it can be suitable to perform a risk analysis.
  \begin{itemize}
    \item Procuring new services?
    \item Outsouring
    \item Organisational changes
  \end{itemize}
\end{frame}

\begin{frame}{How should the process for Risk Analysis look like?}{\insertsubsubsectionhead}
  \footnotesize
  \begin{block}{Already have a process?}
    If the organisation already have a process in place, then this can be used.
    Since it's already well established in the organisation.
  \end{block} 
  \begin{block}{If there isn't a process in place?}
    Some key points to think about when establishing a process for Risk
    Analysis.
    \begin{itemize}
      \item How should the process look like?
      \item What criteria for consequence and probability estimation should be
        used?
      \item What criteria for risk acceptance should be used?
      \end{itemize}
  \end{block}
  \begin{block}{Establish Criterias for Consequence and Probability}
    In IS0 27001:2018 the management are responsible for making a decision
    about the process, criteria for consequence and probability levels, and
    what risk levels should be considered acceptable. 
  \end{block}
\end{frame}
\begin{frame}{Establishing a table for grading consequence}{\insertsubsubsectionhead}
  Create a table for help establishing consequence level connected to each risk.
  For example:
  \begin{itemize}
    \item Financial loss
    \item Negative effect
    \item Infringement
    \item Environmental
    \item Personal Safety
  \end{itemize}
\end{frame}
\begin{frame}{Establishing a table for grading consequence}{\insertsubsubsectionhead}
  \begin{table}
    \tiny
    \caption{Table for helping estimate risk~\cite{msb_metodstod}}
    \begin{tabularx}{\textwidth}{X | X | X | X}
    Consequence & Financial loss & Loss of reputation & Disruption in the
    organisation\\ 
    \toprule

    Severe & Two million SEK and above, or 20\% over budget & Continous negative
    reporting in media or by organised groups in social media. Focused on not
    just individuals but the organisation as a whole.& Disruptions in one or
    several critical organisations that are longer than acceptable.\\

    \midrule
    Considerable & 500 000 - 2 million SEK or 10-20\% over budget & Negative
    reporting in local and national media and in organised social media groups.
    Focused mainly on special instances or actions from a certain individual~(s)
    & Disruption in one critical organisation that are longer than acceptable.
    \\ 
    \midrule
    Moderate & 1 - 500 000 or 5-10 \% over budget & A few dissatisfied
    individuals that express themselves in social medias or a smaller notice in
    local media.& Disruptions in one or several non-critical organisations that
    are longer than acceptable.\\ 
    \midrule
    Negligible & No loss & Little or non negative attention in media or social
    media & Negligable disruption in the organisation\\ 
    \bottomrule
  \end{tabularx}
  \end{table}
\end{frame}

@@

\begin{frame}{Statement of Applicability}{\insertsubsubsectionhead}
  Statement of Applicability (Uttalande om tillämplighet)
  \begin{itemize}
    \item What should be covered in the GAP-analysis?
    \item Select security meassures, either from previous analyses, or pick all.  
    \item Once finished, the organisation have a complete list of security
      meassures that are applicable to the organisation.
  \end{itemize}
\end{frame}
\begin{frame}{Perform the GAP analysis}{\insertsubsubsectionhead}
  Based on the list of applicable security meassures
  \begin{itemize}
    \item Is the security meassure already in place?
    \item Estimate the severity of the security meassure.
    \item Describe the current status of this security meassurement.
  \end{itemize}
\end{frame}
\subsubsection{The Results of the Analysis}
\begin{frame}{The result of the analysis}{\insertsubsubsectionhead}
  \begin{itemize}
    \item The result will act as a base when modelling the information security.
    \item Helps to assess how the organisation handles the information security.
    \item Identifying security measures.
  \end{itemize}

>>>>

\begin{frame}{Tieto incident}
  \begin{itemize}
    \item In 2011 Tieto got a hardware problem in one of their
      datacenters~\cite{Lindkvist2012tdf}. This created disruptions in many of
      the companies customers IT-system. For example: 
      \begin{itemize}
        \item All the pharmacies got problems handing out prescription medicine for a
          time.
        \item The swedish vehicle inspection agency could not report the results
          from the vehicle inspections they did during this time.
        \item A large logistical company could not handle its operation
        \item Social offices in Nacka and Sollentuna were unable to pay child
          support
      \end{itemize}
  \end{itemize}
\end{frame}
\begin{frame}{1177 incident}
  \begin{itemize}
    \item In 2019 a sub-contractor of a sub-contractor for a company that was
      handling the IT-system for three Swedish regions health care guides (1177)
      call line, forgot to secure a backup server that stored the recordings
      from the phone calls~\cite{Ottsjo2019}.
    \item Resulted in around 2.7 million (around 170 000 hours) of recordings of
      confidential phone calls got accessible for the public.
    \item The recordings contains everything from medical symptoms to social
      security numbers.
  \end{itemize}
\end{frame}
\begin{frame}{The Swedish Transport Agency}
  \begin{itemize}
    \item In 2017 it was revealed that sensitive information from the Swedish
      Transport Agency database were made available for non-security screened
      employees abroad~\cite{SvTTransport2017}. 
    \item Information included:
      \begin{itemize}
        \item Driver licences (including picture)
        \item Company secrets
        \item Sensitive information about infrastructure (bridges, subways etc)
        \item Information that can be used to identify armoured cars.
      \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Swedish Public Employment Service}
  \begin{itemize}
    \item In 2019 it was found that the Swedish Public Employment Services
      (arbetsförmedlingen) automatic system for reporting activity reports to
      the Swedish Social Insurance Agency (Försäkringskassan) sent incorrect
      reports~\cite{SvTPublicEmpl2019}.
    \item It was estimated that 10 to 15 percent of the decisions were incorrect.
    \item Resulted in that approximately 70 000 decisions could have been wrong.
  \end{itemize}
\end{frame}

@@

\begin{frame}{Dividing informational assets}{\insertsubsubsectionhead}
  \begin{description}
    \item[Primary] Refers to the main information, such as blue prints, logs or
      contracts.
    \item[Secondary] Refers to resources that are required to access the primary
      informational assets.
  \end{description}
  What will happen if you do not have your program that is required to read the
  closed proprietary format that the information is stored with?
\end{frame}
\begin{frame}{Critical Informational Assets}
  \centering
  \begin{block}{What is a critical informational asset?}
    There are a lot of informational assets in an organisation. It is therefore
    necessary to establish what should be considered a critical asset.
  \end{block}
\end{frame}

@@

\subsubsection{Risk analysis}
\begin{frame}{Overview}{\insertsubsubsectionhead}
  \begin{itemize}
    \item Identifies the information security risks.
    \item Can be performed over generally over the organisation, or on a
      specific object.
    \item The identified risks are then assessed based on probability and
      consequence.
  \end{itemize}
  \begin{block}{What to achieve with a risk analysis}
    \begin{itemize}
      \item What can happen?
      \item How probable is it?
      \item What is the consequences?
    \end{itemize}
    Note that the result of the risk analysis shouldn't be seen as the absolute
    truth.
  \end{block}
\end{frame}

\begin{frame}{When should the Risk Analysis be performed?}{\insertsubsubsectionhead}
  \begin{itemize}
    \item Not necessary to perform the risk analysis after analysing the
      organisation or external and legal requirements.
    \item If it is performed after, then the results from these can be used in
      the risk analysis.
    \item Otherwise, it is possible to rely on the existing knowledge of the
      organisation.
  \end{itemize}
\end{frame}
\begin{frame}{When should the Risk Analysis be performed?}{\insertsubsubsectionhead}
  Examples of when it can be suitable to perform a risk analysis.
  \begin{itemize}
    \item Procuring new services?
    \item Outsourcing
    \item Organisational changes
  \end{itemize}
\end{frame}

\begin{frame}{How should the process for Risk Analysis look like?}{\insertsubsubsectionhead}
  \footnotesize
  \begin{block}{Already have a process?}
    If the organisation already have a process in place, then this can be used.
    Since it's already well established in the organisation.
  \end{block} 
  \begin{block}{If there isn't a process in place?}
    Some key points to think about when establishing a process for Risk
    Analysis.
    \begin{itemize}
      \item How should the process look like?
      \item What criteria for consequence and probability estimation should be
        used?
      \item What criteria for risk acceptance should be used?
      \end{itemize}
  \end{block}
  \begin{block}{Establish Criteria for Consequence and Probability}
    In IS0 27001:2018 the management are responsible for making a decision
    about the process, criteria for consequence and probability levels, and
    what risk levels should be considered acceptable. 
  \end{block}
\end{frame}
\begin{frame}{Establishing a table for grading consequence}{\insertsubsubsectionhead}
  Create a table for help establishing consequence level connected to each risk.
  For example:
  \begin{itemize}
    \item Financial loss
    \item Negative effect
    \item Infringement
    \item Environmental
    \item Personal Safety
  \end{itemize}
\end{frame}
\begin{frame}{Establishing a table for grading consequence}{\insertsubsubsectionhead}
  \begin{table}
    \tiny
    \caption{Table for helping estimate risk~\cite{msb_metodstod}}
    \begin{tabularx}{\textwidth}{X | X | X | X}
    Consequence & Financial loss & Loss of reputation & Disruption in the
    organisation\\ 
    \toprule

    Severe & Two million SEK and above, or 20\% over budget & Continuous negative
    reporting in media or by organised groups in social media. Focused on not
    just individuals but the organisation as a whole.& Disruptions in one or
    several critical organisations that are longer than acceptable.\\

    \midrule
    Considerable & 500 000 - 2 million SEK or 10-20\% over budget & Negative
    reporting in local and national media and in organised social media groups.
    Focused mainly on special instances or actions from a certain individual~(s)
    & Disruption in one critical organisation that are longer than acceptable.
    \\ 
    \midrule
    Moderate & 1 - 500 000 or 5-10 \% over budget & A few dissatisfied
    individuals that express themselves in social media or a smaller notice in
    local media.& Disruptions in one or several non-critical organisations that
    are longer than acceptable.\\ 
    \midrule
    Negligible & No loss & Little or non negative attention in media or social
    media & Negligible disruption in the organisation\\ 
    \bottomrule
  \end{tabularx}
  \end{table}
\end{frame}

@@

\begin{frame}{Statement of Applicability}{\insertsubsubsectionhead}
  Statement of Applicability (Uttalande om tillämplighet)
  \begin{itemize}
    \item What should be covered in the GAP-analysis?
    \item Select security measures, either from previous analyses, or pick all.  
    \item Once finished, the organisation have a complete list of security
      measures that are applicable to the organisation.
  \end{itemize}
\end{frame}
\begin{frame}{Perform the GAP analysis}{\insertsubsubsectionhead}
  Based on the list of applicable security measures
  \begin{itemize}
    \item Is the security measure already in place?
    \item Estimate the severity of the security measure.
    \item Describe the current status of this security measurement.
  \end{itemize}
\end{frame}
\subsubsection{The Results of the Analysis}
\begin{frame}{The result of the analysis}{\insertsubsubsectionhead}
  \begin{itemize}
    \item The result will act as a base when modelling the information security.
    \item Helps to assess how the organisation handles the information security.
    \item Identifying security measures.
  \end{itemize}
